\section{Conclusion}
\label{sec:conclusion}

This work has explored the integration of C++ Standard Parallelism (stdpar) with SYCL through the AdaptiveCpp
compiler, demonstrating how a tightly integrated compiler approach can enable efficient parallel execution on
heterogeneous hardware. By leveraging SYCL as a backend, AdaptiveCpp enables stdpar-based applications to
execute seamlessly across CPUs, GPUs, and other accelerators, maintaining both performance and portability.

\paragraph{Summary of Findings}
The study has shown that AdaptiveCpp’s compiler-based approach provides several key advantages:

\begin{itemize}
    \item \textbf{Tighter integration with the compiler provides significant advantages}, allowing for more efficient
          optimizations compared to purely runtime-based solutions such as NVIDIA’s \texttt{nvc++}.
    \item \textbf{Optimized execution}, reducing kernel launch overhead, improving memory management, and enhancing
          synchronization, leading to more efficient parallel execution.
    \item \textbf{Seamless memory handling}, as USM pooling and optimized buffer-accessor usage significantly reduce
          data transfer overhead and improve execution efficiency.
    \item \textbf{Conditional offloading is a general strategy} that ensures workloads are executed on the most suitable
          hardware, avoiding inefficient GPU offloading for small computations.
\end{itemize}

\paragraph{Implications for the Future of Parallel Computing}
The success of AdaptiveCpp in integrating stdpar with SYCL highlights several broader implications:

\begin{itemize}
    \item \textbf{Lowering the bar for the use of modern hardware accelerators}, making parallel computing more
          accessible to developers without specialized hardware knowledge.
    \item \textbf{Enhancing the C++ ecosystem} by demonstrating how SYCL can serve as a robust backend for standardized
          parallel programming models, fostering greater adoption of heterogeneous computing in modern C++
          development.
    \item \textbf{Encouraging future compiler optimizations} that further improve execution performance and hardware
          adaptability across diverse architectures.
\end{itemize}

While this work has demonstrated the feasibility of using SYCL as a backend for stdpar, additional optimizations
and refinements can further improve execution efficiency and scalability. The following section discusses possible
directions for future research and enhancements.
