\documentclass[manuscript,nonacm]{acmart}

\settopmatter{printacmref=false}


\title{Harnessing the full power of modern hardware accelerators using idomatic C++}

\author{Stefan Butz}
\affiliation{%
    \institution{University of Hagen}
    \city{Hagen}
    \country{Germany}
}
\email{stefan.butz@studium.feruni-hagen.de}

\begin{document}

\begin{abstract}
Abstract is still in construction.
\end{abstract}

\keywords{C++; SYCL; parallelism; CUDA; GPU; stdpar}

\maketitle

\section{Motivation}
% 1 Seiten

% 1. Was ist das Thema? Warum ist das Thema für die Forschung relevant? Welche Forschungsfragen
% werden in der Arbeit beantwortet?
% Einleitung des Themas, Motivation

Modern hardware accelerators, such as GPUs and FPGAs, offer significant computational power.
They are essential for improving the performance of many applications.
However, using these accelerators is challenging because they require specialized knowledge of
their architectures and programming models.

To address this, many libraries and frameworks have been created.
Low-level libraries like CUDA for NVIDIA GPUs and ROCm for AMD GPUs provide detailed control over hardware.
However, they also demand in-depth understanding of hardware-specific programming.
This makes them difficult for developers who are not experts in high-performance computing.

Higher-level frameworks aim to simplify programming for accelerators while maintaining good performance.
OpenCL, for instance, provides a platform-neutral model that works across CPUs, GPUs, and other accelerators.
SYCL builds on OpenCL, offering a modern C++ interface that aligns with standard C++ practices.
A key strength of SYCL is its multi-platform support, allowing developers to write portable code that can
run on a variety of hardware, including GPUs, CPUs, and FPGAs, without modification.
This feature ensures that applications can scale across diverse hardware architectures, providing flexibility
and ease of use.

Despite these advancements, many frameworks still require developers to learn programming models and manage
tasks like device selection and memory transfers.
This adds complexity and slows adoption, especially for those outside specialized computing fields.
Simplifying this process can save time and reduce the learning curve for developers.

Cost is another important factor.
While accelerators improve computational performance, they also require substantial investments in hardware.
Additionally, the time developers spend learning and coding for these systems contributes to project costs.
Simplifying the programming process helps reduce these expenses.
It allows developers to focus on solving domain-specific problems rather than hardware-specific challenges.
Faster project completion also brings business benefits, such as quicker time-to-market and improved competitiveness.

C++ is one of the most widely used programming languages.
It is known for its performance, flexibility, and large ecosystem.
With the introduction of parallel algorithms in C++17, the language has made a major step towards simplifying
parallel programming.
These algorithms allow developers to express parallelism at a high level, avoiding the need to understand the details of
the underlying hardware.

In their research, Aksel Alpay and Vincent Heuveline demonstrated how the SYCL programming model can act as a
backend for C++ standard parallelism.
A key advantage of this approach is that developers can use idiomatic C++ without having to
learn a new programming model. By simply utilizing AdaptiveCpp’s compiler, developers can seamlessly integrate
SYCL into their workflows. This approach ensures that parallelism is implemented in a familiar C++ environment
while still achieving portability and efficiency across various hardware platforms.
This combination of SYCL and C++ standard parallelism creates a unified and accessible approach to modern
parallel programming.

As the foundation of this work, understanding the SYCL programming model is essential.
The next section introduces SYCL, its core principles, and its design, explaining how it provides a modern,
high-level interface for heterogeneous computing while maintaining compatibility with standard C++ practices.

\section{Introduction of the SYCL programming model}
% 4 Seiten
% 2. Grundlagen/Stand der Technik
%     1. Welche Grundlagen sind wichtig, um den Rest der Arbeit zu verstehen?

\subsection{The SYCL programming model}

\subsection{AdaptiveCpp: An SYCL implemenation}

\section{Unifying SYCL and C++ standard parallelism}
% 5 Seiten
% Implementierung/ Beschreibung der Lösungsansätze
%     - Wie können die Forschungsfragen beantwortet werden?
%     - Wie ist das Vorgehen?

\section{Implemenation and performance evaluation}
% 3 Seiten

\section{Conclusion}
% 2 Seiten

% Diskussion der Ergebnisse

\section{outlook}
% 1 Seite
ToDo

% Fazit und Ausblick
\cite{alpay2021adaptivecpp}


\bibliographystyle{acm}
\bibliography{refs} 

\end{document}